A Gaussian process model of human electrocorticographic data

### 2020-02-20 paper/main.tex
did a pass through the discussion

Here we have taken a model-based approach to studying whether high
spatiotemporal resolution activity patterns throughout the human brain may be
explained by a static connectome model that is shared across people and tasks.
Specifically, we trained a model to take in recordings from a subset of brain
locations, and then predicted activity patterns during the same interval, but at
\textit{other} locations that were held out from the model.  Our model, based on
Gaussian process regression, was built on three general hypotheses about the
nature of the correlational structure of neural activity (each of which we
tested).  First, we hypothesized that functional correlations are stable over
time and across tasks.  We found that, although aspects of the patients'
functional correlations were stable across tasks, we achieved better
reconstruction accuracy when we trained the model on within-task data~\citep[we
acknowledge that our general approach could potentially be extended to better
model across-task changes, following][and others]{ColeEtal16, TavoEtal16}.
Second, we hypothesized that some of the correlational structure of people's
brain activity is similar across individuals.  Consistent with this hypothesis,
our model explained the data best when we trained the correlation model using
data from \textit{other} patients-- even when compared to a correlation model
trained on the same patient's data.  Third, we resolved ambiguities in the data
by hypothesizing that neural activity from nearby sources will tend to be
similar, all else being equal.  This hypothesis was supported through our
finding that all of the models we trained that incorporated this spatial
smoothness assumption predicted held-out data well above chance.

One potential limitation of our approach is that it does not provide a natural
means of estimating the precise timing of single-neuron action potentials.
Prior work has shown that gamma band and broadband activity in the LFP may be
used to estimate the firing rates of neurons that underly the population
contributing to the LFP~\citep{MillEtal08, MannEtal09, JacoEtal10b, CronEtal11}.
Because SuperEEG reconstructs LFPs throughout the brain, one could in principle
use broadband power in the reconstructed signals to estimate the corresponding
firing rates (though not the timings of individual action potentials).
Interestingly, our frequency-specific analysis yielded the best reconstruction
accuracies using broadband power.

Beyond providing a means of estimating ongoing activity throughout the brain
using already implanted electrodes, our work also has implications for where to
place the electrodes in the first place. Electrodes are typically implanted to
maximize coverage of suspected epileptogenic tissue.  However, our findings
suggest that this approach could be further optimized.  Specifically, one could
leverage not only the non-invasive recordings taken during an initial monitoring
period (as is currently done routinely), but also recordings collected from
other patients.  We could then ask: given what we learn from other patients'
data (and potentially from the scalp EEG recordings of this new patient), where
should we place a fixed number of electrodes to maximize our ability to map
seizure foci?  As shown in Figures~\ref{fig:informap} and~\intersectmap,
recordings from different locations are differently informative in terms of
reconstructing the spatiotemporal activity patterns throughout the brain.  This
property might be leveraged in decisions about where to surgically implant
electrodes in future patients.

>>>>

Here we have taken a model-based approach to studying whether high
spatiotemporal resolution activity patterns throughout the human brain may be
explained by a static connectome model that is shared across people and tasks.
Specifically, we trained a model to take in recordings from a subset of brain
locations, and then predicted activity patterns during the same interval, but at
\textit{other} locations that were held out from the model.  Our model, based on
Gaussian process regression, was built on three general hypotheses about the
nature of the correlational structure of neural activity (each of which we
tested).  First, we hypothesized that functional correlations are stable over
time and across tasks.  We found that, although aspects of the patients'
functional correlations were stable across tasks, we achieved better
reconstruction accuracy when we trained the model on within-task data.
This suggests that our general approach could be extended to better
model across-task changes, following \cite{ColeEtal16, TavoEtal16} and others.
Second, we hypothesized that some of the correlational structure of people's
brain activity is similar across individuals.  Consistent with this hypothesis,
our model explained the data best when we trained the correlation model using
data from \textit{other} patients-- even when compared to a correlation model
trained on the same patient's data.  Third, we resolved ambiguities in the data
by hypothesizing that neural activity from nearby sources will tend to be
similar, all else being equal.  This hypothesis was supported through our
finding that all of the models we trained that incorporated this spatial
smoothness assumption predicted held-out data well above chance.

One potential limitation of our approach is that it does not provide a natural
means of estimating the precise timing of single-neuron action potentials.
Prior work has shown that gamma band and broadband activity in the LFP may be
used to estimate the firing rates of neurons that underly the population
contributing to the LFP~\citep{MillEtal08, MannEtal09, JacoEtal10b, CronEtal11}.
Because SuperEEG reconstructs LFPs throughout the brain, one could in principle
use broadband power in the reconstructed signals to estimate the corresponding
firing rates (though not the timings of individual action potentials).  We found
that we were able to reconstruct full-brain patterns of broadband power well (Fig.~\ref{fig:frequency}).

Beyond providing a means of estimating ongoing activity throughout the brain
using already implanted electrodes, our work also has implications how to
optimize electrode placements in neurosurgical evaluations. Electrodes are
typically implanted to maximize coverage of suspected epileptogenic tissue.
However, our findings suggest that this approach might be improved.
Specifically, one could leverage not only the non-invasive recordings taken
during an initial monitoring period (as is currently done routinely), but also
recordings collected from \textit{other} patients.  We could then ask: given
what we learn from other patients' data (and potentially from the scalp EEG
recordings of this new patient), where should we place a fixed number of
electrodes to maximize our ability to map seizure foci?  As shown in
Figures~\ref{fig:informap}, \ref{fig:networks}, and~\intersectmap, recordings
from different regions vary in their ability to predict different narrowband and
broadband full-brain activity patterns.

By providing a means of reconstructing full-brain activity patterns, the
SuperEEG approach maps ECoG recordings from different patients into a common
neural space, despite that different patients' electrodes were implanted in
different locations. This feature of our approach enables across-patient ECoG
studies, analogous to across-patient fMRI studies~\citep[e.g.,][]{HaxbEtal01,
NormEtal06, HaxbEtal11}. Whereas the focus of this manuscript is to specifically
evaluate which aspects of neural activity patterns SuperEEG recovers well (or
poorly), in parallel work we are training across-patient classifiers by
leveraging the common neural spaces obtained by applying SuperEEG to
multi-patient data.  For example, we have shown that SuperEEG-derived activity
patterns may be used to accurately predict psychiatric conditions such as
depression~\citep{ScanEtal20}.  Analogous approaches could in principle be used
to develop improved brain-computer interfaces and/or to carry out other analyses
that would benefit from high spatiotemporal resolution full-brain data in
individuals.
